---
title: "JAMES: Joint Anthropometric Measurement and Evaluation Server"
author: "Stef van Buuren"
date: "14-06-2019"
bibliography: [groeidiagrammen.bib]
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
```

## JAMES

The Joint Anthropometric Measurement and Evaluation Server (JAMES) is
an **experimental** online resource for creating growth charts. The
resource can be used by anyone interested in high-quality charts for
monitoring and evaluating childhood growth.

This document explains how to set up communications with JAMES and
provides some guidance in its use.

## Architecture

The growth charts in JAMES are programmed in `R`. JAMES makes these
available through the [OpenCPU](https://www.opencpu.org) system for
scientific computing and reproducuble research. The system allows for
easy integration of growth charts into any `HTTP` compliant client by
means of OpenCPU's [API](https://www.opencpu.org/api.html).

## Access

JAMES is located at url `groeidiagrammen.nl`. The
sections below use `curl` for illustration, but any `HTTP` client will
work.

In order to check whether the JAMES server is running, try generating some random 
numbers by calling `stats::rnorm()`, in a terminal window as follows,

```{bash}
curl https://groeidiagrammen.nl/ocpu/library/stats/R/rnorm/json --data n=5

[2.259, 0.0538, 1.0583, 0.8665, 0.8412]
```

## What JAMES can currently do

1. provides access to 342 high-quality growth charts used by the Dutch youth health care;
2. can accept child data in BDS-format;
3. can upload the data to the server, and then do multiple analyses;
4. can create a personalized site with child's data on growth charts;
5. can create separate growth charts.

## Available growth charts 

Copy the [following url](https://groeidiagrammen.nl/ocpu/lib/james/www/) into your browser to obtain an interactive overview of the available growth charts:

```{bash}
https://groeidiagrammen.nl/ocpu/lib/james/www/
```

There are 342 different charts, each identified with a `chartcode`, 
a 4-7 character code identifying each design. The `list_charts()` function in 
JAMES function can produce a tabular overview. Obtain its help file as

```{bash}
curl https://groeidiagrammen.nl/ocpu/lib/james/man/list_charts/text
```

The list of growth charts can be created by the following request:

```{bash}
curl https://groeidiagrammen.nl/ocpu/lib/james/R/list_charts -d ""

/ocpu/tmp/x045d183eddf2f4/R/.val
/ocpu/tmp/x045d183eddf2f4/R/list_charts
/ocpu/tmp/x045d183eddf2f4/stdout
/ocpu/tmp/x045d183eddf2f4/source
/ocpu/tmp/x045d183eddf2f4/console
/ocpu/tmp/x045d183eddf2f4/info
/ocpu/tmp/x045d183eddf2f4/files/DESCRIPTION
```

which return a list of 8 urls created by `OpenCPU`. The table can be 
download by

```{bash}
curl https://groeidiagrammen.nl/ocpu/tmp/x045d183eddf2f4/R/.val/print

    chartgrp chartcode population    sex design  side language week
1     nl2010      HJAA         HS   male      A front    dutch
2     nl2010      HJAB         HS   male      A  back    dutch
3     nl2010      HJAH         HS   male      A   hgt    dutch
4     nl2010      HJAO         HS   male      A   hdc    dutch
...
```

The chart codes are compatible with the codes used in @talma2010. 

Note that the session key `x045d183eddf2f4` in the above example is
specific to the call, and will be different for each session. Session
keys and their url's on the JAMES server will be removed after 24 
hours.

## Posting child's data to JAMES

Suppose that the child's data are coded according to the specification
[BDS JGZ 3.2.5](https://www.ncj.nl/themadossiers/informatisering/basisdataset/documentatie/?cat=12),
and converted into `JSON` format. The JAMES server contains the
following example:

```{bash}
curl https://groeidiagrammen.nl/ocpu/library/james/testdata/client3.json
```

Save the file
[client3.json](https://groeidiagrammen.nl/ocpu/library/james/testdata/client3.json)
on your local system by

```{bash eval=FALSE}
curl https://groeidiagrammen.nl/ocpu/library/james/testdata/client3.json -O
```

which produces a file called `client3.json` in your work directory. 

For testing purposes, you may change its values (but please keep the
general structure intact) and re-upload it to JAMES for analysis. 

The following `curl` command re-uploads and converts the file
`client3.json` from BDS format into an `R` object of S4 class
`minihealth::individual`:

```{bash, eval = FALSE}
curl -F 'txt=@client3.json' https://groeidiagrammen.nl/ocpu/library/james/R/convert_bds_ind

/ocpu/tmp/x06938035d05dac/R/.val
/ocpu/tmp/x06938035d05dac/R/convert_bds_ind
/ocpu/tmp/x06938035d05dac/stdout
/ocpu/tmp/x06938035d05dac/source
/ocpu/tmp/x06938035d05dac/console
/ocpu/tmp/x06938035d05dac/info
/ocpu/tmp/x06938035d05dac/files/client3.json
/ocpu/tmp/x06938035d05dac/files/DESCRIPTION
```

An print version of the result is stored in 

```{r}
curl https://groeidiagrammen.nl/ocpu/tmp/x06938035d05dac/R/.val/print

An object of class "individual"
Slot "child":
data frame with 0 columns and 0 rows

Slot "time":
data frame with 0 columns and 0 rows

Slot "id":
[1] 0

Slot "name":
[1] "fa308134-069e-49ce-9847-ccdae380ed6f"

Slot "dob":
[1] "11-10-18"

Slot "src":
[1] "1234"

Slot "dnr":
[1] NA

Slot "sex":
[1] "female"

Slot "etn":
[1] "NL"

Slot "edu":
[1] NA

Slot "ga":
[1] 27

Slot "bw":
[1] 990

Slot "twin":
[1] NA

Slot "agem":
[1] NA

Slot "smo":
[1] 0

Slot "hgtm":
[1] 1670

Slot "wgtm":
[1] NA

Slot "hgtf":
[1] 1900

Slot "wgtf":
[1] NA

Slot "bfexc06":
[1] NA

Slot "durbrst":
[1] NA

Slot "hgt":
package: clopus, library: clopus::nl1997 , member: nl1997.fhgtNL 
     age  hgt     hgt.z
1 0.0849 38.0 -7.943717
2 0.1670 43.5 -6.385568

Slot "wgt":
package: clopus, library: clopus::nl1997 , member: nl1997.fwgtNL 
     age  wgt     wgt.z
1 0.0849 1.25 -8.583512
2 0.1670 2.10 -6.581767

Slot "hdc":
package: clopus, library: clopus::nl1997 , member: nl1997.fhdcNL 
     age  hdc     hdc.z
1 0.0849 27.0 -7.548492
2 0.1670 30.5 -6.066642

Slot "bmi":
package: clopus, library: clopus::nl1997 , member: nl1997.fbmiNL 
     age      bmi     bmi.z
1 0.0849  8.65651 -5.719143
2 0.1670 11.09790 -3.766751

Slot "wfh":
No reference
   hgt  wfh wfh.z
1 38.0 1.25    NA
2 43.5 2.10    NA

Slot "bs.hgt":
package: donordata, model: donordata::smocc_bs , member: hgt 
      age      hgt     hgt.z
1  0.0000 37.63782 -6.929621
2  0.0767 38.73560 -7.445715
3  0.1533 42.68033 -6.591987
4  0.2500 46.91763 -5.895737
5  0.3333 50.40139 -5.332599
6  0.5000 55.73412 -4.619819
7  0.6250 58.76424 -4.255843
8  0.7500 61.29885 -4.003211
9  0.9167 63.95774 -3.879229
10 1.1667 67.36804 -3.820196
11 1.5000 71.67160 -3.603667
12 2.0000 77.18910 -3.270500
13 3.0000 90.09290 -1.771377

Slot "bs.wgt":
package: donordata, model: donordata::smocc_bs , member: wgt 
      age        wgt     wgt.z
1  0.0000  0.9056132 -9.006665
2  0.0767  1.2496280 -8.454686
3  0.1533  1.9393741 -6.894393
4  0.2500  2.8894599 -5.392084
5  0.3333  3.5923136 -4.653004
6  0.5000  4.9277774 -3.525314
7  0.6250  5.5292387 -3.306713
8  0.7500  6.1150749 -3.054297
9  0.9167  6.6887088 -2.946767
10 1.1667  7.3461849 -2.920375
11 1.5000  8.2449663 -2.680337
12 2.0000  9.2358047 -2.608811
13 3.0000 10.7343002 -2.773075

Slot "bs.hdc":
package: donordata, model: donordata::smocc_bs , member: hdc 
      age      hdc     hdc.z
1  0.0000 24.74355 -8.185195
2  0.0767 27.30185 -7.188857
3  0.1533 30.11832 -6.148961
4  0.2500 32.12195 -5.860350
5  0.3333 34.30669 -5.121962
6  0.5000 36.13303 -5.082818
7  0.6250 37.72714 -4.571027
8  0.7500 38.47569 -4.519543
9  0.9167 39.62832 -4.234684
10 1.1667 40.34018 -4.247731
11 1.5000 41.79837 -3.710716
12 2.0000 42.51795 -3.634945
13 3.0000 43.21616 -3.808453

Slot "bs.bmi":
package: donordata, model: donordata::smocc_bs , member: bmi 
      age       bmi     bmi.z
1  0.0000  9.191473 -3.230867
2  0.0767  9.055411 -5.031675
3  0.1533 10.752560 -4.023688
4  0.2500 12.188298 -3.201624
5  0.3333 13.524967 -2.264696
6  0.5000 13.881065 -2.343775
7  0.6250 13.829450 -2.527176
8  0.7500 14.487793 -1.969734
9  0.9167 14.597060 -1.853467
10 1.1667 14.419768 -1.878160
11 1.5000 14.525894 -1.568160
12 2.0000 14.376370 -1.485082
13 3.0000 13.781281 -1.783393

Slot "bs.wfh":
Broken stick model not found.
  hgt wfh wfh.z
1  NA  NA    NA
```

The child's data on the server may be used for plotting on growth charts.



A working alternative in `R` (via the `httr` package) is as follows.
Download the example file to your working directory by

```{r downloadexample}
library("httr")
url <- "https://groeidiagrammen.nl/ocpu/library/james/testdata/client3.json"
GET(url, write_disk("client3.json", overwrite = TRUE))
```

Let us now re-upload and convert the local BDS-formatted file
`client3.json` into the JAMES format (defined as S4 class
`minihealth::individual`) by

```{r postfilehttr}
url <- "https://groeidiagrammen.nl"
path <- "ocpu/library/james/R/convert_bds_ind"
dat <- upload_file("client3.json")
resp <- POST(url = url, path = path, body = list(txt = dat))
resp
```

The [jamesclient](https://github.com/stefvanbuuren/jamesclient) package
contains a few utility functions to extract url's for further processing. 
The following code blocks installs the packages, if not yet available.

```{r}
if (!"remotes" %in% rownames(installed.packages())) 
  install.packages("remotes")
if (!"jamesclient" %in% rownames(installed.packages())) 
  remotes::install_github("stefvanbuuren/jamesclient")
```

The return value of the server-side function `convert_bds_ind()` can
now be downloaded as

```{r postfilehttrreturn}
suppressPackageStartupMessages(library("jamesclient"))
url_ret <- get_url(resp, "return")
url_ret
z <- GET(url_ret)
content(z)
```

While this result by itsels is not in a particularly friendly form, it
confirms that the function `convert_bds_ind()` did in fact produce the
intended output. For example, after some searching we may find that
the gestional age (\code{"ga"}) of this child is 27 weeks.

The text version of the help file for `convert_bds_ind()` is at 

```{r postfilehelp}
GET("https://groeidiagrammen.nl/ocpu/library/james/man/convert_bds_ind")
```

Alternatively, paste the following url into your browser, or start a browser 
window from your `R` session to watch the `html` version in your browser by

```{r eval=FALSE}
browseURL("https://groeidiagrammen.nl/ocpu/library/james/man/convert_bds_ind/html")
```

## Requesting a growth chart

### BDS-data as input

The server-side function `james::draw_chart_bds()` generates a growth
chart with the child's BDS-data. It performs three actions: convert
the BDS-data into individual data, select a chart based on the input
data, and draw the chart.

```{r}
path <- "ocpu/library/james/R/draw_chart_bds"
resp2 <- POST(url = url, path = path, body = list(txt = dat))
resp2
```

The produced height chart for this child can be added to a markdown document as 

```{r}
url_svg <- get_url(resp2, "svg")
url_svg
tmp <- tempfile(fileext = ".svg")
GET(url_svg, write_disk(tmp))
knitr::include_graphics(tmp)
```

We may display it in a browser as

```{r eval=FALSE}
browseURL(url_svg)
```

### Data on server as input 

The `draw_chart_ind()` function also draws charts, but takes its input
data from a server-side url that contains the already converted
individual data. Hence, conversion of the BDS-data need to be done
only once. More specifically, `draw_chart_ind()` loads the
`individual` data stored on the server in a previous call, and return
the requested chart. It works as follows:

```{r}
# request preterm chart with all measures for boys, week 25
path <- "ocpu/library/james/R/draw_chart_ind"
resp3 <- POST(url = url, path = path,
              body = list(location = get_url(resp, "location"),
                          chartcode = "PMAAN25",
                          curve_interpolation = TRUE),
              encode = "json")
resp3
```

This growth chart can displayed in the same way as before:

```{r}
tmp <- tempfile(fileext = ".svg")
url_svg <- get_url(resp3, "svg", pad = "?width=10&height=12")
GET(url_svg, write_disk(tmp))
knitr::include_graphics(tmp)
```

Note that the reference chart for height has changed (since we asked
for week 25 instead of week 27), and the child's measurements are
identical.

We can also view it by pasting the url in the browser:

```{r}
url_svg
```

Note that these url's are temporary, and will disappear from the server 
after 24 hours.

## A client in `R`

If `R` is your analysis environment, then you can use the
`jamesclient` package to achieve the same result in an easier way.
The functions `upload_bds()` and `request_chart()` in `jamesclient`
have simple syntax, and automate the steps outlined above. See
<https://github.com/stefvanbuuren/jamesclient> for more detail.

## Next steps

The next step is to extend the system so that individual data can 
be easily plotted and manipulated using a simple app. A first version
(one that does not yet use individual plotting, and that uses 
Dutch growth reference primarily) is on
<https://groeidiagrammen.nl/ocpu/library/james/www/>.

## About

Work in progress. Suggestions and inquiries to Stef van Buuren
(stef.vanbuuren at tno.nl), <https://stefvanbuuren.name>, <https://github.com/stefvanbuuren>.



