---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# jamesclient

<!-- badges: start -->
<!-- badges: end -->

The goal of `jamesclient` is to facilitate interaction for `R` users with the 
**Joint Anthropometric Measurement and Evaluation System (JAMES)**.
JAMES is an **experimental** online resource for creating and analysing
growth charts.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
install.packages("remotes")
remotes::install_github("stefvanbuuren/jamesclient")
```
## Example

The primary functions are:

Function          | Description
----------------- | -------------
`upload_bds()`    | Upload BDS-data, convert to `individual` format
`request_chart()` | Upload BDS-data, draw chart with child data
`request_site()`  | Upload BDS-data, creates a personalised site

### `upload_bds()`

Upload BDS data and create object of class `individual` on server:

```{r example, eval=FALSE}
library(jamesclient)
fn <- file.path(path.package("jamesclient"), "testdata", "client3.json")
r1 <- upload_bds(fn)
browseURL(get_url(r1, "return"))
```

### `request_chart()`

Make a combined upload and automatic chartcode choice:

```{r example2, eval = FALSE}
r2 <- request_chart(fn)
browseURL(get_url(r2, "svg"))
```

### `request_site()`

Combine upload and construction of interactive site:

```{r example3, eval = FALSE}
url <- request_site(fn)
browseURL(url)
```
