---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# jamesclient

<!-- badges: start -->
[![R-CMD-check](https://github.com/growthcharts/jamesclient/workflows/R-CMD-check/badge.svg)](https://github.com/growthcharts/jamesclient/actions)
<!-- badges: end -->

The goal  of `jamesclient` is to facilitate interaction for `R` users with the 
**Joint Automatic Measurement and Evaluation System (JAMES)**.
JAMES is an online resource for creating growth charts and analysing growth curves.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
install.packages("remotes")
remotes::install_github("growthcharts/jamesclient")
```
## Example

The primary functions are:

Function            | Description
------------------- | -------------
`james_post`        | Send POST request to JAMES
`james_get`         | Send GET request to JAMES
`inspect_demodata()`| Upload demo child data and download parsed file

### `james_post()`

Upload BDS data

```{r}
library(jamesclient)
fn <- system.file("extdata", "bds_v3.0", "smocc", "Laura_S.json",
                  package = "jamesdemodata", mustWork = TRUE)
host <- "https://james.groeidiagrammen.nl"
r1 <- james_post(host = host, path = "data/upload/json", txt = fn)
r1$parsed
```

### `james_get()`

Get output from R console (just for checking)

```{r}
r2 <- james_get(host = host, path = paste(r1$session, "console", sep = "/"))
cat(r2$parsed, "\n")
```

For other end points, see <https://james.groeidiagrammen.nl>.

### `inspect_demodata()`

```{r example3}
library(jamesclient)
data <- inspect_demodata(name = "Anne_S", "smocc")
data
```

## Some older functions

Everything can be done with `james_post()` and `james_get()`. The functions below are not needed anymore, and will be deprecated in the future.

### `upload_txt()`

Upload BDS data and create a tibble on the server:

```{r example, eval=FALSE}
library(jamesclient)
fn <- file.path(path.package("jamesclient"), "testdata", "client3.json")
r1 <- upload_txt(fn)
browseURL(get_url(r1, "return"))
```

### `request_chart()`

Make a combined upload and automatic chartcode choice:

```{r example2, eval = FALSE}
r2 <- request_chart(fn, chartcode = "PJAHN27")
browseURL(get_url(r2, "svg"))
```


## Removed functions

Function          | Description                          | Alternative             |
----------------- | -------------------------------------| ----------------------- |
`request_site()`  | Create personalised site             | `james_post()`          |
`upload_bds()`    | Upload and parse data                | `james_post()`          |
